Evolutionary algorithms in a volunteer computing system: performance assessment

### 2016-03-12 PPSN2016/paper.Rnw
Rewriting state of the art
Introducing new factors in human computing. Refs #108.

Please @unintendedbear @mariosky @fergunet take a look at the new approach to the state of the art.

Volunteer computing involves a user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}. However
the dual introduction of JavaScript as a universal language for the
browser and the browser as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here, and whose first version
was described in \cite{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used for either unwitting
\cite{unwitting-ec,boldrin2007distributed} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210}. Many other researchers have
%Pablo: Desell in 2008 is not so recent
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99,DBLP:conf/3pgcic/GuervosMFEL12}. These computing
platforms avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower; this makes gathering new users
immediate and spontaneous, so that they can be used for short
algorithms on call.

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer systems
\cite{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\cite{nogueras2015studying} or including it in a more general study of the
performance of the EA itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\cite{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. On the contrary, in this paper, we will use {\em
  actual} volunteers, anonymous persons that are requested to join an
experiment through social networks.

Some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining,agajaj} found out that the Weibull, log-normal and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles.

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer, that is, a spontaneously created parallel computer that
uses social networks for operations such as gathering new
users. Apol{\'o}nia et al. \cite{apolonia2012enhancing} used the 
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. As in the previous example, a social network
was used to get new network nodes; in the previous case a web page was
used, while Facebook's wall was used here.

In our case, social networks are an integral part of the system and
used to spontaneously obtain users. The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.

>>>>

Volunteer computing involves a user running a program voluntarily
and, as such, has been deployed in many different ways from the
beginning of the Internet, starting with the SETI@home framework for
processing extraterrestrial signals \cite{david-seti:home}. However
the dual introduction of JavaScript as a universal language for the
browser and the browser itself as an ubiquitous web and Internet client has
made this combination the most popular for volunteer computing
frameworks such as the one we are using here, and whose first version
was described in \cite{DBLP:conf/gecco/GuervosG15}.

JavaScript can be used for either unwitting
\cite{unwitting-ec,boldrin2007distributed} or volunteer 
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts
\cite{duda2013distributed,DBLP:journals/corr/abs-0801-1210,EvoStar2014:jsEO}.

In general, those papers go no further than a proof of concept,
attempting to gauge how many users join the effort and also how long
would it take a concrete amount of users to reach the solution. In
general, they do not try to assess the performance of the algorithm
itself in the precise conditions they find in a volunteer computing
system, although papers such as the one by Laredo et
al. \cite{churn08:ijhpsa}, using models, try to find out the resulting
performance when the users are changing continously according to a
Weibull distribution. In the case of Klein and Spector \cite{unwitting-ec}, the algorithm
is actually run on the server, using the browser mainly for fitness
evaluation so no actual contribution to the evaluation of the
evolutionary algorithm itself is made.

There is another factor that should be taken into account: since the
user has control of the browser, there is a limited amount of
interaction with it, namely, the fact that by reloading the webpage he
can apply a kind of hypermutation killing the current population and
generating new individuals some of which will make their way to the
common pool via migration. In that sense, volunteer computing is also
a way of {\em human computation} \cite{quinn2011human}, a concept that
has also been applied to evolutionary algorithms \cite{972056}, in
this case extensively and with all operators. It is quite difficult to
find out how many times this happens and what is their effect on the
overall algorithm, but this only reaffirms the fact that there is a
long list of issues with volunteer evolutionary computation, and that
the volunteer him or herself is at the center of many of them.

In this paper we will make contributions towards the design of
efficient volunteer evolutionary algorithms by studying the effect of
several design decisions on performance. In the next Section we will
present the general framework and the initial setup.
